---
title: "species_table"
date: "January 8 2025"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{species_table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bowen.biodiversity.webapp)
library(here)
library(stringr)
library(dplyr)
library(googlesheets4)
library(googledrive)
library(foreach)
library(sf)
```

# Environmental Spatial Data Layers 
Run this RMarkdown script to download environmental spatial data layers to correct location within repository. Files are stored on Google Drive. 

```{r download}
# Download data to location available to Shiny application. 
download_data_path <- here::here("inst", "app", "data")

#### Islands Trust Spatial Data Layers ####
# https://islandstrust.bc.ca/mapping-resources/mapping/bowen/ 

# Cadastral (Property) Shapefile from Islands Trust
bowen_island_properties_zipfile_path <- here::here(download_data_path, "bowen_island_properties.zip")
bowen_island_properties_extracted_path <- here::here(download_data_path, "bowen_island_properties")
if(!dir.exists(bowen_island_properties_extracted_path)) {
  googledrive::drive_download(file = "https://drive.google.com/file/d/16ys45qQdcnyJb1hUoG7Mi7cxUgTctZDI/view?usp=drive_link",
                              path = bowen_island_properties_zipfile_path,
                              overwrite = T)
  unzip(zipfile = bowen_island_properties_zipfile_path,
        exdir = bowen_island_properties_extracted_path)
  file.remove(bowen_island_properties_zipfile_path)
}

# Protected Areas Shapefile from Islands Trust
bowen_island_protectedareas_zipfile_path <- here::here(download_data_path, "bowen_island_protectedareas.zip")
bowen_island_protectedareas_extracted_path <- here::here(download_data_path, "bowen_island_protectedareas")
if(!dir.exists(bowen_island_protectedareas_extracted_path)) {
  googledrive::drive_download(file = "https://drive.google.com/file/d/1DXpqHi7RdfM3hImS3qTzzWH1dwj-dmu8/view?usp=drive_link",
                              path = bowen_island_protectedareas_zipfile_path,
                              overwrite = T)
  unzip(zipfile = bowen_island_protectedareas_zipfile_path,
        exdir = bowen_island_protectedareas_extracted_path)
  file.remove(bowen_island_protectedareas_zipfile_path)
}

# Zoning Shapefile from Islands Trust
bowen_island_zoning_zipfile_path <- here::here(download_data_path, "bowen_island_zoning.zip")
bowen_island_zoning_extracted_path <- here::here(download_data_path, "bowen_island_zoning")
if(!dir.exists(bowen_island_zoning_extracted_path)) {
  googledrive::drive_download(file = "https://drive.google.com/file/d/1RFqBN38KRZ_VEA3733JYWSa_aF_P-YBh/view?usp=drive_link",
                              path = bowen_island_zoning_zipfile_path,
                              overwrite = T)
  unzip(zipfile = bowen_island_zoning_zipfile_path,
        exdir = bowen_island_zoning_extracted_path)
  file.remove(bowen_island_zoning_zipfile_path)
}
#### Bowen Island Spatial Data Layers from John Dowler ####
bowen_island_covenants_path <- here::here(download_data_path, "bowen_island_covenants.gpkg")
if(!file.exists(bowen_island_covenants_path)) {
  googledrive::drive_download(file = "https://drive.google.com/file/d/1VY9DG0uosIWDnHviGwb0t_FvLIEkqfO7/view?usp=drive_link",
                              path = bowen_island_covenants_path,
                              overwrite = T)
}
#### Metro Vancouver Sensitive Ecosystems Inventory ####
# https://arcg.is/TnfXL 
mvsei_path <- here::here(download_data_path, "bowen_island_mvsei.gpkg")
bowen_island_sei_path <- here::here("inst", "app", "data", "bowen_island_sei.gpkg")
if(!file.exists(bowen_island_sei_path)) {
  googledrive::drive_download(file = "https://drive.google.com/file/d/1lmH7UA0kBBBwnfaJDd9uIuVK0Aolwt_c/view?usp=drive_link",
                              path = mvsei_path,
                              overwrite = T)
  mvsei <- sf::st_read(mvsei_path)
  bowen_island_boundary <- sf::st_read(here::here("data-raw", "bowen_boundary")) %>%
    sf::st_transform(st_crs(mvsei))
  bowen_island_sei <- sf::st_intersection(mvsei, bowen_island_boundary)
  st_write(bowen_island_sei, bowen_island_sei_path)
  file.remove(mvsei_path)
  
  #### Translate MV SEI ecosystem class codes to description using values from data docs ####
  mvsei_SECl_codes <- data.frame(
    code = unique(mvsei$SECl_1)
  )
  mvsei_SECl_codes$description <- c(
    "Young Forest (small)",
    "Riparian",
    "Lost - Young Forest (small)",
    "Mature Forest",
    "Young Forest",
    "Lost - Riparian",
    "Lost - Mature Forest",
    "Lost - Young Forest",
    "Lost - Old Field",
    "Lost - Wetland",
    "Freshwater",
    "Old Field",
    "Wetland",
    "Estuarine",
    "Intertidal",
    "Old Forest",
    "Herbaceous",
    "Woodland",
    "Sparsely Vegetated",
    "Lost - Herbaceous",
    "Non SE or ME",
    "Lost - Estuarine",
    "Alpine",
    "Lost - Freshwater",
    "Lost - Old Forest",
    "Lost - Woodland",
    "Lost - Intertidal",
    "Lost - Sparsely Vegetated"
  )
  write.csv(mvsei_SECl_codes, 
            here::here(download_data_path, 
                       "mvsei_SECl_codes.csv"), 
            row.names = F)
}

#### Existing Disturbance ####
# TODO: replace with Google Drive Download
disturbed_download_path <- "/media/jay/2TB_Drive/2024_Bowen_Biodiversity_Project/human_footprint_30m_v2.tif"
bc_disturbed_path <- here::here(download_data_path, "bc_human_footprint_30m.tif")
bowen_island_undisturbed_path <- here::here(download_data_path, "bowen_island_env_rasters", "bowen_island_normal_inverse_human_footprint.tif")
if(!file.exists(bowen_island_undisturbed_path)) {
  file.copy(disturbed_download_path, bc_disturbed_path)
  # Load raster
  disturbed <- terra::rast(bc_disturbed_path)
  # Resample raster to one of SDMs clipped to Bowen Island
  ex_sdm <- terra::rast(here::here("inst/app/rasters/2024_12_11_SDM_100m_Bowen_Island/Amphibian_mask/Ambystoma.gracile_BowenIsland_100m.tif"))
  disturbed_resamp <- terra::resample(disturbed, ex_sdm)
  # Normalize from 0 to 1
  disturbed_normal <- disturbed_resamp %>% normalize()
  # Invert so that high values are low disturbance and vice versa
  undisturbed <- disturbed_normal %>% invert()
  # Write Bowen Island inverse and normalized human footprint raster
  terra::writeRaster(undisturbed, 
                     bowen_island_undisturbed_path, 
                     overwrite=TRUE)
  # Clean up enormous human footprint file
  file.remove(bc_disturbed_path)
}
```

# Load layers that have been downloaded
```{r}
#### Load layers with terra
# bowen_island_properties <- terra::vect(bowen_island_properties_extracted_path)
# bowen_island_protectedareas <- terra::vect(bowen_island_protectedareas_extracted_path)
# bowen_island_zoning <- terra::vect(bowen_island_zoning_extracted_path)
# bowen_island_covnenants <- terra::vect(bowen_island_covenants_path)
# bowen_island_mvsei <- terra::vect(bowen_island_mvsei_path)

#### Load layers with sf
bowen_island_properties <- sf::st_read(bowen_island_properties_extracted_path)
bowen_island_protectedareas <- sf::st_read(bowen_island_protectedareas_extracted_path)
bowen_island_zoning <- sf::st_read(bowen_island_zoning_extracted_path)
bowen_island_covenants <- sf::st_read(bowen_island_covenants_path)
bowen_island_sei <- sf::st_read(bowen_island_sei_path)
```


# Extract ecosystem types from MV SEI and turn into new raster layers
```{r}
mvsei_SECl_codes <- read.csv(here::here("inst", "app", "data", "mvsei_SECl_codes.csv"))

#### Decide which MV SEI classes to turn into habitat rasters to include into Zonation run ####
# include_classes <- c(
#   "Mature Forest",
#   "Old Forest",
#   "Woodland",
#   "Riparian",
#   "Wetland",
#   "Freshwater",
#   "Estuarine",
#   "Intertidal",
#   "Herbaceous"
# )


# Create a list of sf objects, each one to become a layer
# Select SEI by SECl and SEsubcl 

# SEI Cl and subcl categories 
SE_cat <- bowen_island_sei[,c("SECl_1", "SEsubcl_1")] %>%
  st_drop_geometry() %>% 
  unique()
rownames(SE_cat) <- paste0(SE_cat$SECl_1, "_", SE_cat$SEsubcl_1)

# SEsubcl descriptions
bsei_SEsubcl_codes <- SE_cat$SEsubcl_1 %>% 
  unique() %>%
  data.frame(SEsubcl = .) %>% 
  na.omit()
bsei_SEsubcl_codes$description <- c(
  "mixed conifer and broadleaf trees (< 75% coniferous and < 75% broadleaf composition)",
  "conifer dominated (> 75% stand composition)",
  "coastal herbaceous: criteria as for ‘hb’ but influenced by proximity to ocean; windswept shoreline and slopes; > 20% vegetation of grasses, herbs, mosses and lichens.",
  "fringe: narrow linear communities along open water bodies (rivers, streams, lakes and ponds) where there is no floodplain. see Appendix I for mapping guidelines.",
  "swamp: wooded wetlands dominated by 25% or more cover of flood-tolerant trees or shrubs. Swamps are characterized by periodic flooding and nearly permanent sub-surface waterflow through mixtures of mineral and organic materials, swamps are high in nutrient, mineral and oxygen content.",
  "rock outcrop: exposed bedrock, usually at the top of knolls or on portions of steeper slopes.",
  "vegetated shoreline: low-lying rocky shoreline, soil pockets in rock cracks and crevices; salt-tolerant vegetation, generally with < 20% vegetation cover.",
  "mudflats: non-vegetated mudflats or with varying amounts of algae.",
  "gully: watercourse is within a steep sided V-shaped gully or ravine; generally only minimal area of flooding but gully is important due to proximity to water and sensitive due to steeper slopes.",
  "fen: underlain by sedge or brown moss peat, fens are closely related to bogs. In addition to rainfall, fens receive mineral and nutrient-enriched water from upslope drainage or groundwater. Thus a broader range of plants, including shrubs and small trees, is able to grow.",
  "shallow water: wetlands characterized by water less than 2m in depth in mid-summer; transition between deep water bodies and other wetland ecosystems (i.e. bogs, swamps, fens, etc.); often with vegetation rooted below the water surface.",
  "beaches: well- to sparsely-vegetated or non-vegetated beaches and shorelines.",
  "pond: naturally occurring, small body of open water, greater than 2m deep and generally up to 8 ha, with little to no floating vegetation; shallower water than a lake.",
 "broadleaf dominated (> 75% of stand composition), any size.",
 "eelgrass: intertidal & shallow subtidal eelgrass beds.",
 "lake: naturally occurring, static body of open water greater than 2m deep and generally greater than 8 ha, with little to no floating vegetation; deeper water than a pond.",
 "bog: acidic, nutrient-poor wetlands that characteristically support peat-mosses and ericaceous shrubs such as Labrador tea and bog-rosemary. Being generally isolated from mineral rich groundwater or surface water, their primary source of water and nutrients is from rainfall.",
 "Non SE, ME or YS ecosystem type."
)

layers_to_rast <- foreach(i = 1:nrow(SE_cat),
                          .final = function(i) {
                            setNames(i, rownames(SE_cat))
                          }) %do% {
  SECl <- SE_cat[i,"SECl_1"]
  SEsubcl <- SE_cat[i,"SEsubcl_1"]
  
  select_sei <- bowen_island_sei %>%
    filter(SECl_1 == SECl) %>%
    filter(SEsubcl_1 == SEsubcl)
}

# Raster generation code
foreach(layer_name = names(layers_to_rast)) %do% {
  layer_to_rast <- layers_to_rast[[layer_name]]
  # Turn selected SEI into raster
  sei_rast <- bowen_sf_to_rast(layer_to_rast)
  # Save raster for Zonation runs later
  if(!dir.exists(here::here("inst", "app", "data", "bowen_island_sei_all_cl_subcl"))) {
      dir.create(here::here("inst", "app", "data", "bowen_island_sei_all_cl_subcl"))
  }
  rast_path <- here::here("inst", "app", "data", "bowen_island_sei_all_cl_subcl", paste0(layer_name, "_raster.tiff"))
  if(!file.exists(rast_path)) {
      terra::writeRaster(sei_rast,
                     filename = rast_path,
                     overwrite = T)
  }
}

# Create plots to review habitat rasters
habitat_rasters <- list.files(here::here("inst", "app", "data", "bowen_island_sei_all_cl_subcl"))
foreach(habitat_raster = habitat_rasters) %do% {
  sei_cl_subcl <- habitat_raster %>%
    gsub("_raster.tiff", "", .) %>%
    stringr::str_split("_")
  sei_cl_desc <- mvsei_SECl_codes[mvsei_SECl_codes$code == sei_cl_subcl[[1]][1], "description"]
  sei_subcl_desc <- bsei_SEsubcl_codes[bsei_SEsubcl_codes$SEsubcl == sei_cl_subcl[[1]][2], "description"]
  
  habitat_rast <- terra::rast(here::here("inst", "app", "data", "bowen_island_env_rasters", "bowen_island_sei", habitat_raster))
  # Plot
  bowen_plot <- bowen_map(raster_layer = habitat_rast,
                          title = paste0(sei_cl_desc, " Class: ", sei_cl_subcl[[1]][1], " Subclass: ", sei_cl_subcl[[1]][2]),
                          subtitle = sei_subcl_desc,
                          caption = paste0("Map created on ", Sys.time(), " by Jay Matsushiba (jmatsush@sfu.ca) in the Palen Lab."),
                          legend_label = "Presence")
  # Save plot
  output_path <- here::here("vignettes/figures/individual_habitat_layers")
  dir.create(output_path)
  ggplot2::ggsave(
    file = paste0(output_path, "/", sei_cl_subcl[[1]][1], "_", sei_cl_subcl[[1]][2], ".png"),
    bowen_plot,
    device = ragg::agg_png,
    width = 9, height = 12, units = "in", res = 300
  )
}

#### TAKE TWO #####
# Young Forest
bowen_island_sei_YF <- bowen_island_sei[bowen_island_sei$SECl_1 == "YF",]
# Young Forest (small)
bowen_island_sei_YS <- bowen_island_sei[bowen_island_sei$SECl_1 == "YS",]
# Woodland
bowen_island_sei_WD <- bowen_island_sei[bowen_island_sei$SECl_1 == "WD",]
# Herbaceous
bowen_island_sei_HB <- bowen_island_sei[bowen_island_sei$SECl_1 == "HB",]
# Mature Forest
bowen_island_sei_MF <- bowen_island_sei[bowen_island_sei$SECl_1 == "MF",]
# Riparian
bowen_island_sei_RI <- bowen_island_sei[bowen_island_sei$SECl_1 == "RI",]
# Wetland
bowen_island_sei_WN <- bowen_island_sei[bowen_island_sei$SECl_1 == "WN",]

# Intertidal
bowen_island_sei_IT <- bowen_island_sei[bowen_island_sei$SECl_1 == "IT",]
## Mudflats
bowen_island_sei_IT_mf <- bowen_island_sei_IT %>% filter(SEsubcl == "mf")
## Eelgrass
bowen_island_sei_IT_el <- bowen_island_sei_IT %>% filter(SEsubcl == "el")
## Beaches
bowen_island_sei_IT_bs <- bowen_island_sei_IT %>% filter(SEsubcl == "bs")

# Old Forest
bowen_island_sei_OF <- bowen_island_sei[bowen_island_sei$SECl_1 == "OF",]

layers_to_rast <- list(
  YF = bowen_island_sei_YF,
  YS = bowen_island_sei_YS,
  WD = bowen_island_sei_WD,
  HB = bowen_island_sei_HB,
  MF = bowen_island_sei_MF,
  RI = bowen_island_sei_RI,
  WN = bowen_island_sei_WN,
  IT_mf = bowen_island_sei_IT_mf,
  IT_el = bowen_island_sei_IT_el,
  IT_bs = bowen_island_sei_IT_bs,
  OF = bowen_island_sei_OF
)

# Raster generation code
foreach(layer_name = names(layers_to_rast)) %do% {
  layer_to_rast <- layers_to_rast[[layer_name]]
  # Turn selected SEI into raster
  sei_rast <- bowen_sf_to_rast(layer_to_rast)
  # Save raster for Zonation runs later
  if(!dir.exists(here::here("inst", "app", "data", "bowen_island_env_rasters"))) {
      dir.create(here::here("inst", "app", "data", "bowen_island_env_rasters"))
  }
  rast_path <- here::here("inst", "app", "data", "bowen_island_env_rasters", paste0(layer_name, "_raster.tiff"))
  if(!file.exists(rast_path)) {
      terra::writeRaster(sei_rast,
                     filename = rast_path,
                     overwrite = T)
  }
}
```

# Other environmental layers / features
```{r}
ex_sdm <- terra::rast(here::here("inst/app/rasters/2024_12_11_SDM_100m_Bowen_Island/Amphibian_mask/Ambystoma.gracile_BowenIsland_100m.tif"))

bowen_island_ogma_path <- here::here(download_data_path, "bowen_island_env_rasters", "bowen_island_ogma.tif")
#### OGMA (Old Growth Management Areas) ####
# Extract from protectedareas
if(!file.exists(bowen_island_ogma_path)) {
  bowen_island_ogma <- bowen_island_protectedareas %>%
    filter(NAME == "Old Growth Management Area") %>%
    terra::vect() %>%
    terra::project(ex_sdm)
  bowen_island_ogma_rast <- terra::rasterize(bowen_island_ogma,
                                             ex_sdm) 
  terra::writeRaster(bowen_island_ogma_rast, 
                     bowen_island_ogma_path, 
                     overwrite=T)
}
```

# Create environmental layers table
```{r table}
# select only environmental rasters
bowen_island_env_rasters_dir <- "inst/app/data/bowen_island_env_rasters"
# relative paths within project
rasters_paths <- list.files(here::here(bowen_island_env_rasters_dir),
                            recursive = T)
layers_df <- data.frame(full_path = paste(bowen_island_env_rasters_dir, rasters_paths, sep = "/"),
                        filename = rasters_paths,
                        weights = 1.0,
                        layer_name = "")

# This sheet was uploaded to Google Sheets
# WARNING: DO NOT RUN OR WILL OVERWRITE MANUAL INPUT TO GOOGLE SHEETS
# write_sheet(layers_df,
#             "https://docs.google.com/spreadsheets/d/1S7U53ukf74y3YaCDZUUSCh2Rv_wpbdZ4veH-BotoQKo/edit?gid=0#gid=0",
#             "environmental")
# write.csv(layers_df, here::here("inst/app/data/bowen_island_env_rasters.csv"))
```

# Read in from Google Sheets 
```{r}
layers_googlesheet <- read_sheet("https://docs.google.com/spreadsheets/d/1S7U53ukf74y3YaCDZUUSCh2Rv_wpbdZ4veH-BotoQKo/edit?gid=0#gid=0",
                                 "environmental")
```


