---
title: "zonation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zonation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Bowen Island Biodiversity Prioritization - Zonation5
https://github.com/zonationteam/Zonation5 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#### Palen Lab Packages ####
library(bowen.biodiversity.webapp)
# remotes::install_github("Palen-Lab/sdmcsr")
library(sdmcsr) # SDM congruence functions with iNaturalist observations
# remotes::install_github("Palen-Lab/rinat") # Fork by Jay Matsushiba
library(rinat) 

#### Other Packages ####
library(here)
library(dplyr)
library(stringr)
library(foreach)
library(sf)
library(terra)
library(tidyterra)
library(readxl)
library(googlesheets4)
```
## Loading Datasets

### Notes on Data Management
Raw data (temp files) into `data-raw` and added to `data-raw/.gitignore` Save finished outputs (as used by the application) into `inst/extdata` - Convention is to save files that are not `.rda` into this folder (<https://rstudio.github.io/r-manuals/r-exts/Creating-R-packages.html#data-in-packages>)

### Species Distribution Models
```{r}
#### SPECIES DISTRIBUTION MODELS ####
# Contents of "data-raw/species_distribution_models/" downloaded from Cumulative effects - SFU Teams Sharepoint 
# https://1sfu.sharepoint.com/:f:/r/teams/Cumulativeeffects-SFUTeams9/Shared%20Documents/BCH_project_backup/SPECIES?csf=1&web=1&e=FpNkgc 
# Directory Path: BCH_project_backup/SPECIES/
# Paper: https://www.nature.com/articles/s41598-020-64501-7#MOESM1
# See Wendy Palen (SFU) for access 
#### PROCESSING SDM ####
# Get list of available SDMs
sdm_folder <- here("data-raw/species_distribution_models")
sdm_folder_tif <- list.files(sdm_folder, recursive = T, pattern = "*.tif$") # Get only .tif files
# Correspondence with Viorel Popescu confirms that _NAs_ should be SDMs with rock, ice, large lakes layers masked
sdm_folder_tif_NAs <- sdm_folder_tif[grepl("_NAs_", sdm_folder_tif)] # Get only .tif with "_NAs_" in filename
# Defining CRS to make sure all SDMs have the same
sdm_crs <- "+proj=aea +lat_0=45 +lon_0=-126 +lat_1=50 +lat_2=58.5 +x_0=1000000 +y_0=0 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
# Bowen Island Administrative Boundary (from Metro Vancouver)
bowen_island_admin <- vect(here("data-raw/bowen_boundary")) %>%
  project(sdm_crs)
# Directory to save SDMs
sdm_output_dir <- here("inst/extdata/species_distribution_models_bowen")
if(!dir.exists(sdm_output_dir)) {
  dir.create(sdm_output_dir)
}
# Create Bowen Island mask using DSM
sample_sdm <- rast(here(sdm_folder, sdm_folder_tif_NAs[1])) %>%
  terra::project(y = sdm_crs) %>%
  terra::crop(bowen_island_admin, snap = "out") %>%
  terra::disagg(fact = 4)
dsm <- rast(here("data-raw/bowen_dsm/chm.tif"))
dsm_smooth <- dsm %>% 
  focal(w=9, fun=mean, na.policy="only", na.rm=T) %>% 
  terra::project(y = sdm_crs) %>%
  terra::resample(sample_sdm, method = "average")
m <- c(-1, 35, 1) %>% matrix(ncol = 3, byrow=T)
bowen_island_mask <- dsm_smooth %>% 
  classify(m)
# Weights for terra::focal() function that provides moving window average
# - Smoothing algorithm 
# - Use this to fill some empty cells on Bowen Island with weighted mean of the neighbouring cells.
weights <- matrix(c(0, 0, 1, 1, 1, 0, 0,
                    0, 1, 1, 2, 1, 1, 0,
                    1, 1, 3, 3, 3, 1, 1,
                    1, 2, 3, 5, 3, 2, 1,
                    1, 1, 3, 3, 3, 1, 1,
                    0, 1, 1, 2, 1, 1, 0,
                    0, 0, 1, 1, 1, 0, 0
), nrow=7)
#### PREPARE SDMs TO BOWEN ISLAND ####
# Loop through SDMs
foreach(sdm_tif = sdm_folder_tif_NAs) %do% {
  sdm_path <- here(sdm_folder, sdm_tif) 
  sdm_output_path <- here(sdm_output_dir, basename(sdm_tif))
  sdm <- rast(sdm_path)
  # Normalize SDMs that range from 0 to 1000
  sdm_minmax <- terra::minmax(sdm)
  if(sdm_minmax[2] > 1) {
    sdm <- sdm / 1000
  }
  sdm_output <- sdm %>%
    #### Reproject SDM ####
    terra::project(y = sdm_crs) %>%
    #### Crop SDM ####
    # Figured out why the SDMs were cropped incorrectly
    # The default for cropping a raster by a polygon is to snap = "near"
    # This means that the polygon extent is snapped to the closest raster
    # We need snap = "out" to make sure the full extent of the polygon is covered by the output raster
    terra::crop(bowen_island_admin, snap = "out") %>%
    #### Extrapolate for Southern Parts of Bowen Island
    # - fills the NA values missing in south part of Bowen Island
    terra::focal(w = weights,
                 fun = "mean",
                 na.policy = "only") %>%
    #### Downsamples ####
    # - from 400 m to 100 m resolution
    terra::disagg(fact = 4) %>%
    #### Smooths the raster ####
    terra::focal(w = weights,
                 fun = "mean",
                 na.rm = T,
                 na.policy = "omit") %>%
    terra::mask(bowen_island_mask)
  
  writeRaster(sdm_output, sdm_output_path, overwrite = T)
}
# Create species list from available SDMs
species_list <- sdm_folder_tif_NAs %>%
  basename() %>%
  stringr::str_extract("^\\S+\\.[a-zA-Z]+_") %>%
  stringr::str_remove("_") %>%
  stringr::str_replace("\\.", " ")
# Create taxon list 
taxon_list <- sdm_folder_tif_NAs %>%
  dirname() %>%
  str_remove("_mask")
# Get iNaturalist and SDM congruence
sdm_inat <- foreach(i = 1:length(species_list), .combine = "rbind") %do% {
  # Progress Message 
  print(paste0("Currently working on species ", i, " out of ", length(species_list),". This is ", species_list[i], "."))
  # Query iNaturalist by species name and Bowen Island Administrative Boundaries
  rinat_results_sf <- tryCatch({
    Sys.sleep(3) # Need to slow down queries to prevent requests being blocked
    rinat::get_inat_obs_sf(taxon_name = species_list[i],
                           area = st_as_sf(bowen_island_admin),
                           maxresults = 10000)
  }, error = function(cond) {
    NA
  })
  # Get number of iNaturalist observations
  if(!is.data.frame(rinat_results_sf)) {
    inat_n_obs <- 0
  } else {
    inat_n_obs <- nrow(rinat_results_sf)
  }
  # Get Bowen Island SDM
  sdm <- paste0(str_replace(species_list[i], " ", "."), "_NAs_pu_mask.tif") %>%
    here(sdm_output_dir, .) %>%
    rast()
  # Get highest value in SDM
  sdm_max <- sdm %>% 
    minmax() %>% 
    max()
  
  # PREPARE OUTPUT ROW
  # 1. scientific species name
  # 2. highest SDM cell value within area 
  # 3. observation count from iNaturalist
  result <- data.frame(species = species_list[i],
                       sdm_max_value = sdm_max,
                       inat_n_obs = inat_n_obs)
}
sdm_inat$taxon_group <- taxon_list
#### Query eBird for birds ####
# Number of eBird observations by bird species on Bowen Island
# - Results from Natasha Beauregard (Summer 2024 USRA)
bowen_ebird <- read.csv(here("data-raw/bowen_ebird/final_bird.csv")) %>%
  select(c("scientific_name", "n_ebird"))
# Combine eBird and iNaturalist counts
bowen_SDM_iNat_ebird <- merge(sdm_inat, 
                              bowen_ebird, 
                              by.x = "species", 
                              by.y = "scientific_name",
                              all.x = TRUE) %>%
  rename(n_obs_inat = inat_n_obs,
         n_obs_ebird = n_ebird) %>%
  select(
    species,
    sdm_max_value,
    taxon_group,
    n_obs_inat,
    n_obs_ebird
  )
# Query Pottinger Report (2005) for small mammals
bowen_mammals <- read_xlsx(here("data-raw/pottinger_report/Appendix 2 - Expected Occurrence of Wildlife Species In The Cape Roger Curtis Area.xlsx")) %>%
  filter(`Documented on Bowen` == "y" | `Documented on Bowen` == "?" | `Documented on Bowen` == "One or more spp" )
# Include Species if they meet any of the following criteria:
# - eBird observations > 0
# - iNaturalist observations > 0
# - included in Pottinger Report 
include_threshold <- 0.5 # Min SDM cell value somewhere on Bowen Island for including species into focal list
bowen_SDM_obs_focal_df <- bowen_SDM_iNat_ebird %>%
  mutate(include = case_when(
    (n_obs_inat > 0 | n_obs_ebird > 0 ) & sdm_max_value > include_threshold ~ "y",
    species %in% bowen_mammals$`Scientific name` ~ "y")
  )
# Create list of species names to include
bowen_focal_list <- bowen_SDM_obs_focal_df %>%
  filter(include == "y") %>%
  select(species) %>%
  unlist() %>%
  unname() 
# Get filepath for each SDM
bowen_SDM_iNat_ebird$filepath <- lapply(bowen_SDM_iNat_ebird$species, function(x) {paste0(str_replace(x, " ", "."), "_NAs_pu_mask.tif")})
# Add columns that can be adjusted manually
bowen_SDM_iNat_ebird$prov_status <- ""
bowen_SDM_iNat_ebird$fed_status <- ""
bowen_SDM_iNat_ebird$IUCN_status <- ""
bowen_SDM_iNat_ebird$weights <- 1.0
# Dataframe for focal species on Bowen Island
bowen_sdm_focal <- bowen_SDM_iNat_ebird %>%
  filter(species %in% bowen_focal_list)
#### UPLOAD ####
# This sheet was uploaded to Google Sheets
# WARNING: DO NOT RUN OR WILL OVERWRITE MANUAL INPUT TO GOOGLE SHEETS
# write_sheet(bowen_sdm_focal,
#             "https://docs.google.com/spreadsheets/d/1ehO30w4i7EDafilNyRpq3sgc86fm8g_TmYMI4ql3qms/edit?usp=sharing",
#             "species")
# write.csv(bowen_sdm_focal, here("inst/extdata/species_distribution_models_focal_bowen_info.csv"))
```
#### Creating Species Information Spreadsheet 
This spreadsheet contains all of the information on species name, taxon group, threatened status, weights to use for Zonation. 

Visit this link to see the spreadsheet hosted on Google Drive.
https://docs.google.com/spreadsheets/d/1ehO30w4i7EDafilNyRpq3sgc86fm8g_TmYMI4ql3qms/edit?usp=sharing
```{r}
bowen_sdm_focal <- read.csv(here("inst/extdata/species_distribution_models_focal_bowen_info.csv"))
#### REVISION ####
# - This section necessary, since Google Sheet was uploaded at earlier time
# - Changes since made to the creation of Bowen SDM Focal
# - Michael Tylo (USRA Spring 2025) did the work in the Google Sheet
# Manually look up species and correct taxonomy, add conservation status
sdm_googlesheet <- read_sheet("https://docs.google.com/spreadsheets/d/1ehO30w4i7EDafilNyRpq3sgc86fm8g_TmYMI4ql3qms/edit?usp=sharing") %>%
  select(!c("full_path", "filename"))
# Drop duplicate columns, keep columns for n eBird and iNaturalist observations, max SDM values
bowen_sdm_focal_sel <- bowen_sdm_focal %>%
  select(!c("taxon_group", "prov_status", "fed_status", "IUCN_status", "weights"))
# Merge
bowen_sdm_status <- merge(bowen_sdm_focal_sel, sdm_googlesheet, by.x = "species", by.y = "sci_name", all.x = T) 
save(bowen_sdm_focal, file = here("inst/extdata/bowen_sdm_focal.rda"))
```
### Metro Vancouver Sensitive Ecosystems Inventory
```{r}
#### Metro Vancouver Sensitive Ecosystems Inventory ####
# Contents of "data-raw/metrovancouver_sensitive_ecosystem_inventory/" downloaded from MetroVancouver Open Data Catalogue
# GeoPackage download: https://arcg.is/TnfXL 
# Reports download: https://metrovancouver.org/services/regional-planning/sensitive-ecosystem-inventory-mapping-app
```
### Human Disturbance
```{r}
#### Human Disturbance #### 
# Contents of "data-raw/human_disturbance/" received from Miguel Arias Patino (marias@unbc.ca)
# https://1sfu.sharepoint.com/:f:/r/teams/Cumulativeeffects-SFUTeams9/Shared%20Documents/General/BC%20Hydro%202025%20IRP/Environmental%20Layers/Human%20Footprint?csf=1&web=1&e=hUYDs7 
# Thesis (methods): https://unbc.arcabc.ca/islandora/object/unbc%3A59107
# Paper (methods): https://doi.org/10.1016/j.ecolind.2024.112407 
# Paper (methods): https://doi.org/10.1139/facets-2021-0063  

#### Private Lands ####

#### Protected Areas #### 

```
## Data Processing 
Creating individual layers
```{r}
#### Species Distribution Models ####

#### Metro Vancouver Sensitive Ecosystems Inventory ####

#### Human Disturbance #### 

#### Private Lands ####

#### Protected Areas #### 

```
## Zonation 
Running zonation with all inputs

## Output
