---
title: "zonation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{zonation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# Bowen Island Biodiversity Prioritization - Zonation5
https://github.com/zonationteam/Zonation5 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(bowen.biodiversity.webapp)
library(here)
library(dplyr)
library(stringr)
library(sf)
library(terra)
library(tidyterra)
```
## Loading Datasets

### Species Distribution Models

### Metro Vancouver Sensitive Ecosystems Inventory

### Human Disturbance

### Private Lands

### Protected Areas

### Notes on Data Management
Raw data (temp files) into `data-raw` and added to `data-raw/.gitignore` Save finished outputs (as used by the application) into `inst/extdata` - Convention is to save files that are not `.rda` into this folder (<https://rstudio.github.io/r-manuals/r-exts/Creating-R-packages.html#data-in-packages>)

```{r}
#### Species Distribution Models ####
# Contents of "data-raw/species-distribution-models" downloaded from Cumulative effects - SFU Teams Sharepoint 
# https://1sfu.sharepoint.com/:f:/r/teams/Cumulativeeffects-SFUTeams9/Shared%20Documents/BCH_project_backup/SPECIES?csf=1&web=1&e=FpNkgc 
# Directory Path: BCH_project_backup/SPECIES/
# Paper: https://www.nature.com/articles/s41598-020-64501-7#MOESM1
# See Wendy Palen (SFU) for access 

#### Metro Vancouver Sensitive Ecosystems Inventory ####
# Contents of "data-raw/metrovancouver_sensitive_ecosystem_inventory" downloaded from MetroVancouver Open Data Catalogue
# GeoPackage download: https://arcg.is/TnfXL 
# Reports download: https://metrovancouver.org/services/regional-planning/sensitive-ecosystem-inventory-mapping-app

#### Human Disturbance #### 
# Contents of "data-raw/human_disturbance" received from Miguel Arias Patino (marias@unbc.ca)
# https://1sfu.sharepoint.com/:f:/r/teams/Cumulativeeffects-SFUTeams9/Shared%20Documents/General/BC%20Hydro%202025%20IRP/Environmental%20Layers/Human%20Footprint?csf=1&web=1&e=hUYDs7 
# Thesis (methods): https://unbc.arcabc.ca/islandora/object/unbc%3A59107
# Paper (methods): https://doi.org/10.1016/j.ecolind.2024.112407 
# Paper (methods): https://doi.org/10.1139/facets-2021-0063  

#### Private Lands ####

#### Protected Areas #### 

```
## Data Processing 
Creating individual layers
```{r}
#### Species Distribution Models ####

#### Metro Vancouver Sensitive Ecosystems Inventory ####

#### Human Disturbance #### 

#### Private Lands ####

#### Protected Areas #### 

```
## Zonation 
Running zonation with all inputs

## Output
